---
title: "Homework 6"
author: "Safiya Sirota"
date:  2021-12-04
output: 
  github_document:
    toc: TRUE
---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(modelr)
library(mgcv)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

__Importing and cleaning the data:__

```{r}
birthwt_df <- read_csv("data/birthweight.csv") %>% 
  mutate(
    babysex = as.factor(babysex),
    frace = as.factor(frace),
    malform = as.factor(malform),
    mrace = as.factor(mrace)
  )

map(birthwt_df, ~sum(is.na(.)))
map(birthwt_df, ~sum(is.character(.)))
```

__Fitting a linear model__

```{r}
map(birthwt_df, ~cor.test(as.numeric(.), birthwt_df$bwt))
```

sex, head, delwt, fincome, frace, gaweeks, mheight, momage, mrace, ppbmi, smoken, and wtgain significantly correlated

```{r}
fit <- lm(bwt ~ babysex + bhead + delwt + frace + gaweeks + mheight + momage + mrace + ppbmi + smoken + wtgain, data = birthwt_df)
broom::tidy(fit)
```

We can get rid of the insignificant variables

```{r}
fit2 <- lm(bwt ~ babysex + bhead + gaweeks + mrace + smoken, data = birthwt_df)
broom::tidy(fit2)
```

Graph residuals

```{r}
modelr::add_residuals(birthwt_df, fit2) %>% 
  ggplot(aes(x = bwt, y = resid)) +
  geom_point()
#not totally linear

modelr::add_residuals(birthwt_df, fit2) %>% 
  ggplot(aes(x = resid)) +
  geom_density()

#add predictions
```


## Problem 2

__Importing data:__

```{r}
weather_df <-
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

r^2
log(β^0∗β^1)

Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities. Plot the distribution of your estimates, and describe these in words. Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for r^2 and log(β^0∗β^1). Note: broom::glance() is helpful for extracting r^2 from a fitted regression, and broom::tidy() (with some additional wrangling) should help in computing log(β^0∗β^1).

```{r}
weather_df %>% 
  bootstrap(n = 5000, id = "strap_number") %>% 
  mutate(
    models = map(.x = strap, ~lm(tmax ~ tmin, data = .x)),
    results = map(models, broom::tidy)
  ) %>% 
  select(strap_number, results) %>% 
  unnest(results)
```

